---
title: "Penguin flipper lengths anova"
author: "Sean Denny"
date: "11/7/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r}
library(tidyverse)
library(car)
library(onewaytests)
penguins <- read_csv("penguins.csv")
```

Our question: Is there a significant difference in mean flipper lengths for female chinstrap, gentoo, and adelie penguins near Palmer Archipelago?

Maybe I can use a one-way ANOVA... but there are assumptions tied to that:
-Independent observations
-Equal variance (Leven's test)
-Normality (histograms/qqplots)

```{r}
counts <- with(penguins, table(Species, Sex))

penguins2 <- penguins %>%
  rename(flipper_length = `Flipper Length (mm)`) %>% 
  mutate(
    common_name = 
      case_when(
        Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "chinstrap",
        Species == "Gentoo penguin (Pygoscelis papua)" ~ "gentoo",
        Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "adelie"
      )
) %>% 
  select(-Species) %>% 
  filter(Sex == "FEMALE")
```

Exploratory graphs and Leven's test for equal variances

```{r}
hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~common_name, scale = "free")
hists

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~ common_name, scale = "free")
qqs

# Based on qq-plots and histograms, observations forreach group appear normally distributed Also, n >  30 for all groups, so mean will be normally distibuted by CLT.

```

Evaluate variances and perform a Leven's Test for equal variances.

```{r, message = FALSE}
#If the largest sample variances is < 4x greater than the smallest sample variance, then these are "close enough" to consdier variances equal.

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )
#variances

# The largest variance (33.1) is deiniftely not more than 4x greater than the smaller variance (15.2).

# Leven's Test
# levenTest function in the 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)
#levene_penguin

#H0: No difference in variances (variances are equal)
#HA: Variances are NOT equal

#There is no significant differences in variances across species.
```

One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}

#H0: Means across all groups are equal
#HA: At least two group means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
penguin_aov
summary(penguin_aov)

#Post-hoc testing by Tukey's HSD
#H0: No significant difference in pairwise means
#HA: There is a significant difference in pairwise means

penguin_post_hoc <- TukeyHSD(penguin_aov)
penguin_post_hoc

#aov.test function in the 'onewaytests' packages DOES allow for inline references
```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova; F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).